---
title: "Bioacoustic Vignette"
author: "Reese Hotten-Somers"
date: "2023-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Intro to Bioacoustics:
What is bioacoustics?

Who uses it?

How it's implemented?

What it can be used for and why it is important?

## Preliminaries
Install packages: {bioacoustics}, {warbleR}, {randomForest}, {Rraven}

## Objectives


## Load in Packages
```{r, prelims}
library(warbleR)
library(bioacoustics)
library(tools)
library(randomForest)



```

# Importing Data from the internet
Warblr comes with the function query_xc which allows you to seamlessly import audio data from xeno-canto[https://xeno-canto.org/484847].
```{r, importing}
#pulling all Strepera versicolor calls without downloading locally
df1 = query_xc(qword ='Strepera versicolor type:call cnt:"Australia"', download = FALSE) #todo: include explanation for parameters of query_xc

#Filter by type   
df1 = df1[df1$Vocalization_type=="call",]
df1 = df1[df1$Quality=="A",]
df1 = df1[1:9,]
View(df1)

```

# Importing Your Own Data
However, you can also import your own MP3
```{r, importing_mp3}

```

# Reading Your file
```{r, readf}
query_xc(X = df1, download = TRUE, path = getwd())
```

